A Fully Abstract Game Semantics for Countable Nondeterminism

### 2018-01-30 cones/paper/cones.tex

The second observation is that Lemma \ref{lem:soundness-lemma} can just as easily be applied to PCF contexts and its game semantics model $\G_{inn}$.  
There really is nothing to this observation: since any PCF context may be interpreted in Idealized Algol, then Lemma \ref{lem:soundness-lemma} applies to it.
So we have narrowed \ref{lem:soundness-lemma} down to a result about $\G_{inn}$.  
The only other change we need to make is to switch our base observable type from $\com$ to $\nat$, since PCF has no command type.  
The version with $\nat$ instead of $\com$ can be proved in exactly the same way.

\begin{lemma}
  Let $C[-]\from\nat$ be a PCF context with a hole of type $\nat$.
  Let $\sigma\from\bN\to\bN$ be the denotation of $\lambda n.C[n]$ in $\G_{inn}^\oc$.  
  Then $C[\wn]\converges n$ if and only if there exists some $\s\in\sigma$ such that $\s\vert_\bN=qn$.  
  Moreover, $C[\wn]\mustconverge$ if and only if $\sigma$ is winning.
\end{lemma}

>>>>

We want a version of Lemma \ref{lem:soundness-lemma} that can be applied to PCF contexts and the game semantics model $\G_{inn}$ of PCF.  
Clearly, the lemma as currently stated does not hold in $\G_{inn}$, since the strategies $\sigma$ that we define are not innocent in general.

Instead, we need to look at the combinatorial structure of the strategies.  
We observe that if $\sigma_M\from\oc\bN\to\bC$ is a strategy, then there exists a deterministic strategy $\sigma\from\oc\bN$ such that $qa\in T_{\sigma;\sigma_M}$ if and only if there exists some $\s\in T_{\sigma_M}$ such that $\s\vert_{\bC}=qa$.  
In addition, there exists a deterministic strategy $\sigma\from\oc\bN$ such that $D_{\sigma;\sigma_M}\ne\emptyset$ if and only if $\sigma_M$ is not \emph{winning}; i.e., either $\sigma_M$ is not total or there exists some $\s\in T_{\sigma_M}$ with an infinite tail of moves in $\oc\bN$.  
We recover the following version of Lemma \ref{lem:soundness-lemma} which now applies to the interpretation of PCF in $\G_{inn}$:
\begin{lemma}
  Let $M=C[\wn]$ be a term of type $\com$.  
  Write $\sigma_M\from\oc\bN\to\bC$ for the denotation of the term $\lambda n.C[n]$.  
  \begin{itemize}
    \item $M\converges\skipp$ if and only if there exists $\s\in \sigma_M$ such that $\s\vert_{\bC}=qa$.  
    \item $M\mustconverge$ if and only if $\sigma_M$ is \emph{winning}.
  \end{itemize}
  \label{lem:pcf-soundness-lemma}
\end{lemma}
